### 2018-02-09 Sections/QTCP_protocol_stack.tex
QOS update

\comment{Summary here. Shift bulk to QOS chapter.}

In classical networking theory, error detection is an important element of networking protocols. Communication links may be unreliable, or subject to external noise, which users must be able to detect so as to guarantee the quality of their data.

@@

In the quantum context the elegant notion of checksums is complicated by the fact that calculating a hash function of a quantum state would necessarily entangle the state with the hash function. This would have the undesired effect of causing measurement of the checksum to collapse the quantum state of the data, thereby altering it in an uncontrollable way.

@@

QEC codes have been described for protecting against universal error models, including:
\begin{itemize}
\item Pauli errors: encompassing dephasing, depolarising and amplitude damping \comment{Is A.D a Pauli channel?} error channels.
\item Qubit loss: particularly useful in the optical context where photons are easily lost.
\item Gate failure\index{Gate failure}: essential when gates are non-deterministic, necessarily the case when dealing with maximally entangling two-qubit photonic quantum gates.
\end{itemize}

The first described QEC\index{Quantum error correction (QEC)} code was the 3-qubit code\index{3-qubit code} by \cite{bib:Shor95} for redundantly encoding a single logical qubit into three encoded qubits, allowing the detection and correction of at most a single bit-flip error between the encoded qubits. This protocol is shown in Alg.~\ref{alg:three_QEC}. By switching into a Hadamard-rotated basis, the same code could equivalently correct against at most a single phase-flip error (since \mbox{$\hat{H}\hat{X}\hat{H}=\hat{Z}$}). By concatenating these two codes a 9-qubit code\index{9-qubit code} correcting against a single depolarising error may be implemented.

Since these original ideas into QEC\index{Quantum error correction (QEC)}, countless further codes have been described, able to correct against larger numbers of errors, with far more efficient resource overheads. Currently, topological codes (Sec.~\ref{sec:topol_codes}) are considered the most favourable codes in terms of their elegance, simplicity, high error-correction thresholds, and efficiency in resource overhead. Some such topological codes are particularly enticing, as they can be readily prepared from cluster states (Sec.~\ref{sec:CSQC}), which lend themselves naturally to certain networking methods, such as distributed state preparation.

\begin{table}[!htb]
\fbox{\parbox{0.965\columnwidth}{\texttt{
function ThreeQubitCode($\ket\psi$):
\begin{enumerate}
\item Using two CNOT gates, redundantly encode the logical single-qubit state,
\begin{align}
\ket\psi=\alpha\ket{0}+\beta\ket{1},
\end{align}
into the 3-qubit state,
\begin{align}
\ket\psi_R &= \hat{\text{CNOT}}_{1,2}\hat{\text{CNOT}}_{1,3}\ket\psi\ket{00} \nonumber \\
&= \alpha\ket{000}+\beta\ket{111}.
\end{align}
\item Independently apply bit-flip channels $\mathcal{E}_X$ to each of the three encoded qubits.
\item If exactly one bit-flip operation was applied in total, the three possible erroneous encoded states are,
\begin{align}
\ket\psi_1 &= \hat{X}_1 \ket\psi_R = \alpha\ket{001}+\beta\ket{110}, \nonumber \\
\ket\psi_2 &= \hat{X}_2 \ket\psi_R = \alpha\ket{010}+\beta\ket{101}, \nonumber \\
\ket\psi_3 &= \hat{X}_3 \ket\psi_R = \alpha\ket{100}+\beta\ket{011}.
\end{align}
\item Determine the parity of each of the three pairs of encoded qubits.
\item Assuming at most a single bit-flip operation has occurred on the encoded state, the three parity outcomes uniquely determine which encoded qubit the bit-flip was applied to.
\item Apply classically-controlled bit-flip recovery operations, $\mathcal{R}$, to correct the encoded state, recovering $\ket\psi_R$.
\item Apply the inverse of the encoding operation to recover $\ket\psi$.
\item $\Box$
\end{enumerate}}
\begin{align}
\Qcircuit @C=1.3em @R=.6em {
  & \lstick{\ket{\psi}} & \ctrl{2} & \gate{\mathcal{E}_X}  & \qw & \qw              & \ctrl{3}  & \qw       & \ctrl{4} & \qw & \multigate{2}{\ \mathcal{R}\ } & \qw \\
  & \lstick{\ket{0}}    & \targ    & \gate{\mathcal{E}_X}  & \qw & \qw              & \ctrl{2}  & \ctrl{2}  & \qw      & \qw & \ghost{\ \mathcal{R}\ } \qw & \qw \\
  & \lstick{\ket{0}}    & \targ    & \gate{\mathcal{E}_X}  & \qw & \qw              & \qw       & \ctrl{2}  & \ctrl{3} & \qw & \ghost{\ \mathcal{R}\ } \qw & \qw \\
  &          &          &          & & \lstick{\ket{0}} & \targ \qw & \qw       & \qw      & \meter & \control \cw \cwx \\
  &          &          &          & & \lstick{\ket{0}} & \qw       & \targ \qw & \qw      & \meter & \control \cw \cwx \\
  &          &          &          & & \lstick{\ket{0}} & \qw       & \qw       & \targ    & \meter & \control \cw \cwx
} \nonumber
\end{align}
}}
\caption{3-qubit code for protecting against at most a single logical bit-flip error. The doubly-controlled CNOT gates represent parity measurements, \mbox{$n_3=n_1\oplus n_2$}, where $n_i$ represents the value of the $i$th qubit. In a Hadamard-rotated basis, the same circuit may be employed to protect against a single phase-flip error. And by concatenating the two we obtain a 9-qubit code protecting against a single depolarising error (i.e joint bit-flip/phase-flip), which is a universal single-qubit error model. \comment{Move to QOS section}} \label{alg:three_QEC}
\end{table}

Different error correcting codes have different error correcting power, and different resource overheads, which must be taken into consideration. The appropriate choice of code will largely be determined by the final \textsc{Services \& Applications} layer. For some simple quantum communications protocols, simple error correction (or even no error correction) may suffice. A full-fledged quantum computation on the other hand will require error rates within a fault-tolerance threshold.

The field of fault-tolerance is already extremely well developed and could be applied directly to QTCP as an abstraction layer directly below the \textsc{Services \& Application} layer of the protocol stack -- the \textsc{QoS} layer -- abstracting it away from the end-user. The transport layer protocol for applying it to the QTCP stack is shown in Alg.~\ref{alg:qos}.

\begin{table}[!htb]
\fbox{\parbox{0.965\columnwidth}{\texttt{ 
function QoS(Packets):
\begin{enumerate}
    \item qec = Packets.ApplyQEC()
    \item if(qec.success = true) \{
    \setlength{\itemindent}{0.2in}
    \item message.Receiver.Notify(qec.data)
    \setlength{\itemindent}{0in}
    \item \} else \{
    \setlength{\itemindent}{0.2in}
    \item message.Receiver.Notify(\textsc{Failure})
    \setlength{\itemindent}{0in}
    \item \}
    \item $\Box$
\end{enumerate}}}}
\caption{QoS algorithm based on any appropriate existing QEC code.} \label{alg:qos}
\end{table}

>>>>

In classical networking theory, error detection and correction is an important element of networking protocols. Communication links may be unreliable, or subject to external noise, which users must be able to detect so as to guarantee the quality of their data.

@@

In the quantum context the elegant notion of checksums is complicated by the fact that calculating a hash function of a quantum state would necessarily entangle the state with the output hash. This would have the undesired effect of causing measurement of the checksum to collapse the quantum state of the data, thereby altering it in an uncontrollable way.

@@

Thus, enforcing QoS in fundamentally different for quantum data packets than for classical ones, requiring entirely different techniques. Given how large a field this has become in its own right, we dedicate Sec.~\ref{sec:QOS_chap} entirely to protocols suited to the implementation of quantum QoS requirements in quantum networks.

### 2018-02-09 Sections/protocols_for_the_quantum_internet.tex
QOS update

Quantum key distribution (QKD) is a relatively mature technology with already several commercial systems being available off-the-shelf \footnote{\cite{??? example companies}}. The main drawback is thus the ability to perform long-distance transmission of photons. For these reasons, current QKD networks have been limited to relatively small distances, on the order of $\sim100$km, including in Austria, Switzerland, Japan, USA, and China \cite{bib:lo2014secure}. The longest distance QKD network that is currently planned is the Beijing to Shanghai QKD link spanning a distance of 2000km. This involves 32 trusted nodes\index{Trusted nodes} to break the total distance into shorter segments to convert the quantum information to classical information.

>>>>

Quantum key distribution (QKD) is a relatively mature technology with already several commercial systems being available off-the-shelf\footnote{\cite{??? example companies}}. The main drawback is thus the ability to perform long-distance transmission of photons. For these reasons, current QKD networks have been limited to relatively small distances, on the order of $\sim100$km, including in Austria, Switzerland, Japan, USA, and China \cite{bib:lo2014secure}. The longest distance QKD network that is currently planned is the Beijing to Shanghai QKD link spanning a distance of 2,000km. This involves 32 trusted nodes\index{Trusted nodes} to break the total distance into shorter segments to convert the quantum information to classical information.

### 2018-02-09 Sections/quality_of_service.tex
QOS update

\section{Quality of service (QoS)}\index{Quantum error correction (QEC)}\index{Quality of service}

@@

@@

@@

@@

@@

A single-qubit $Z$-rotation of angle $\theta$ on the Bloch sphere\index{Bloch sphere} is given by,
\begin{align}
	\hat{Z}(\theta) &= e^{i\frac{\theta}{2}\hat{Z}}\nonumber \\
	&= \left(\begin{matrix}
  e^{i\frac{\theta}{2}} & 1 \\
  0 & e^{-i\frac{\theta}{2}}
\end{matrix}\right).
\end{align}
where $\hat{Z}$ is the usual Pauli phase-flip operator\index{Pauli operators},
\begin{align}
\hat{Z}=\ket{0}\bra{0}-\ket{1}\bra{1}.	
\end{align}
This operates on the physical basis states as,
\begin{align}
	\hat{Z}(\theta) \ket{0} &\to e^{i\frac{\theta}{2}}\ket{0}, \nonumber \\
	\hat{Z}(\theta) \ket{1} &\to e^{-i\frac{\theta}{2}}\ket{1}.
\end{align}

Now we employ the encoding for logical basis states,
\begin{align}
\ket{0}_L &\equiv \ket{0}_1\otimes\ket{1}_2,\nonumber \\
\ket{1}_L &\equiv \ket{1}_1\otimes\ket{0}_2,
\end{align}
Note that both logical basis states are invariant under a common $Z$-rotation (the other two physical basis states, $\ket{0}_1\otimes\ket{0}_2$ and $\ket{1}_1\otimes\ket{1}_2$ do not observe this property),
\begin{align}
	\hat{Z}_1(\theta)\hat{Z}_2(\theta)\ket{0}_L &= \ket{0}_L,\nonumber\\
	\hat{Z}_1(\theta)\hat{Z}_2(\theta)\ket{1}_L &= \ket{1}_L.
\end{align}
Thus, when acting on an arbitrary linear combination of these basis states (i.e a logical qubit),
\begin{align}
\ket\psi_L = \alpha \ket{0}_L + \beta\ket{1}_L,
\end{align}
via linearity the logical qubit must also be invariant under the collective error,
\begin{align}
	\hat{Z}_1(\theta)\hat{Z}_2(\theta)\ket\psi_L = \ket\psi_L.
\end{align}
This type of DFS encoding therefore protects a logical qubit against arbitrary correlated $Z$-rotations.

The same principle can be logically extended to many other correlated error models. For example, operating in a rotated basis (under a Hadamard transform), one could similarly protect against correlated $X$-rotations using the encoding,
\begin{align}
\ket{0}_L &\equiv \ket{-}\otimes\ket{+},\nonumber \\
\ket{1}_L &\equiv \ket{+}\otimes\ket{-},
\end{align}
where \mbox{$\ket\pm=\frac{1}{\sqrt{2}}(\ket{0}\pm\ket{1})$}.

This idea, although very simple, is very powerful, since it is a completely passive form of error correction, requiring no syndrome measurements, feedforward or correction operations. It also arises quite naturally in some systems where external fields act roughly uniformly across the physical qubits within a system.

@@

Let us illustrate how this can be achieved using a simple example, whereby a single-qubit system couples with an environment which introduces a slow, unknown phase evolution. Discretising time, we can write this phase evolution as a Pauli $Z$-rotation on the Bloch sphere\index{Bloch sphere},
\begin{align}
	\hat{Z}(\theta) &= e^{i\frac{\theta}{2}\hat{Z}}\nonumber \\
	&= \left(\begin{matrix}
  e^{i\frac{\theta}{2}} & 1 \\
  0 & e^{-i\frac{\theta}{2}}
\end{matrix}\right),
\end{align}
for some unknown, but small $\theta$. Next we observe that, quite obviously, a $Z$-rotation of angle $\theta$ can be trivially undone by applying another $Z$-rotation of angle $-\theta$, since,
\begin{align}
\hat{Z}(-\theta)\hat{Z}(\theta) = \hat{I}.	
\end{align}

@@

Now if we proceed for two time-steps, once bit-flipped and another not, we have,
\begin{align}
\hat{Z}(\theta)\cdot[\hat{X}\hat{Z}(\theta)\hat{X}] = \hat{I},
\end{align}
and the unknown phase-rotation has been eliminated.

Proceeding over the course of a lengthy evolution, we will implement the dynamical decoupling by continuously applying bit-flips to our qubit, such that in conclusion there are no accumulated unknown phase-evolutions. The control sequence for implementing this is sometimes referred to as `bang-bang' control\index{Bang-bang control}, since we are repeatedly implementing bit-flips at a fast rate.

@@

Because the evolution implemented by the computation is constantly changing the state of our computational qubit, the change in trajectory introduced by the dephasing error model is also varying over time. But for the phase cancellation to work, the forward and backward erroneous steps ($\hat{Z}(\pm\theta)$) must be along the same path -- i.e they must all couple-up. This mandates that the bang-bang bit-flip control sequence be executed at a rate very fast compared to the rate of change in the phase-error -- if we fail to implement them fast enough, erroneous steps in the forward and backward directions will not pair up exactly, leaving some residual accumulated errors.

Dynamical decoupling extends to all manner of error models, beyond the simple $Z$-error model presented above. They are therefore a very powerful tool in error correction. However, unfortunately they are only naturally suited to continuous evolutions governed by Hamiltonians of the form shown in (Eq.~\ref{eq:dyn_dec_ham}), not to the more common discretised models such as the circuit and cluster state models.

>>>>

\section{Quality of service (QoS)}\index{Quantum error correction (QEC)}\index{Quality of service}\label{sec:QOS_chap}

@@

Different error correcting codes have different error correcting power, and different resource overheads, which must be taken into consideration. The appropriate choice of code will largely be determined by the final \textsc{Services \& Applications} layer. For some simple quantum communications protocols, simple error correction (or even no error correction) may suffice. A full-fledged quantum computation on the other hand will require error rates within a fault-tolerance threshold\index{Fault-tolerance threshold}.

@@

The first described QEC\index{Quantum error correction (QEC)} code was the 3-qubit code\index{3-qubit code} by \cite{bib:Shor95} for redundantly encoding a single logical qubit into three encoded qubits, allowing the detection and correction of at most a single bit-flip error between the encoded qubits. This protocol is shown in Alg.~\ref{alg:three_QEC}. By switching into a Hadamard-rotated basis, the same code could equivalently correct against at most a single phase-flip error (since \mbox{$\hat{H}\hat{X}\hat{H}=\hat{Z}$}).

\begin{table}[!htb]
\fbox{\parbox{0.965\columnwidth}{\texttt{
function ThreeQubitCode($\ket\psi$):
\begin{enumerate}
\item Using two CNOT gates, redundantly encode the logical single-qubit state,
\begin{align}
\ket\psi=\alpha\ket{0}+\beta\ket{1},
\end{align}
into the 3-qubit state,
\begin{align}
\ket\psi_R &= \hat{\text{CNOT}}_{1,2}\hat{\text{CNOT}}_{1,3}\ket\psi\ket{00} \nonumber \\
&= \alpha\ket{000}+\beta\ket{111}.
\end{align}
\item Independently apply bit-flip channels $\mathcal{E}_X$ to each of the three encoded qubits.
\item If exactly one bit-flip operation was applied in total, the three possible erroneous encoded states are,
\begin{align}
\ket\psi_1 &= \hat{X}_1 \ket\psi_R = \alpha\ket{001}+\beta\ket{110}, \nonumber \\
\ket\psi_2 &= \hat{X}_2 \ket\psi_R = \alpha\ket{010}+\beta\ket{101}, \nonumber \\
\ket\psi_3 &= \hat{X}_3 \ket\psi_R = \alpha\ket{100}+\beta\ket{011}.
\end{align}
\item Determine the parity of each of the three pairs of encoded qubits.
\item Assuming at most a single bit-flip operation has occurred on the encoded state, the three parity outcomes uniquely determine which encoded qubit the bit-flip was applied to.
\item Apply classically-controlled bit-flip recovery operations, $\mathcal{R}$, to correct the encoded state, recovering $\ket\psi_R$.
\item Apply the inverse of the encoding operation to recover $\ket\psi$.
\item $\Box$
\end{enumerate}}
\begin{align}
\Qcircuit @C=1.3em @R=.6em {
  & \lstick{\ket{\psi}} & \ctrl{2} & \gate{\mathcal{E}_X}  & \qw & \qw              & \ctrl{3}  & \qw       & \ctrl{4} & \qw & \multigate{2}{\ \mathcal{R}\ } & \qw \\
  & \lstick{\ket{0}}    & \targ    & \gate{\mathcal{E}_X}  & \qw & \qw              & \ctrl{2}  & \ctrl{2}  & \qw      & \qw & \ghost{\ \mathcal{R}\ } \qw & \qw \\
  & \lstick{\ket{0}}    & \targ    & \gate{\mathcal{E}_X}  & \qw & \qw              & \qw       & \ctrl{2}  & \ctrl{3} & \qw & \ghost{\ \mathcal{R}\ } \qw & \qw \\
  &          &          &          & & \lstick{\ket{0}} & \targ \qw & \qw       & \qw      & \meter & \control \cw \cwx \\
  &          &          &          & & \lstick{\ket{0}} & \qw       & \targ \qw & \qw      & \meter & \control \cw \cwx \\
  &          &          &          & & \lstick{\ket{0}} & \qw       & \qw       & \targ    & \meter & \control \cw \cwx
} \nonumber
\end{align}
}}
\caption{3-qubit code for protecting against at most a single logical bit-flip error. The doubly-controlled CNOT gates represent parity measurements, \mbox{$n_3=n_1\oplus n_2$}, where $n_i$ represents the value of the $i$th qubit. In a Hadamard-rotated basis, the same circuit may be employed to protect against a single phase-flip error. And by concatenating the two we obtain a 9-qubit code protecting against a single depolarising error (i.e joint bit-flip/phase-flip), which is a universal single-qubit error model. \comment{Move to QOS section}} \label{alg:three_QEC}
\end{table}

@@

By concatenating these two codes a 9-qubit code\index{9-qubit code} correcting against a single depolarising error may be implemented.

@@

Currently, topological codes (Sec.~\ref{sec:topol_codes}) are considered the most favourable codes in terms of their elegance, simplicity, high error-correction thresholds, and efficiency in resource overhead. Some such topological codes are particularly enticing, as they can be readily prepared from cluster states (Sec.~\ref{sec:CSQC}), which lend themselves naturally to certain networking methods, such as distributed state preparation.

@@

A single-qubit $Z$-rotation of angle $\theta$ on the Bloch sphere\index{Bloch sphere} is given by,
\begin{align}
	\hat{Z}(\theta) = e^{i\frac{\theta}{2}\hat{Z}} = \left(\begin{matrix}
  e^{i\frac{\theta}{2}} & 0 \\
  0 & e^{-i\frac{\theta}{2}}
\end{matrix}\right).
\end{align}
where $\hat{Z}$ is the usual Pauli phase-flip operator\index{Pauli operators} (Sec.~\ref{sec:circuit_model}). This operates on the physical basis states as,
\begin{align}
	\hat{Z}(\theta) \ket{0} &\to e^{i\frac{\theta}{2}}\ket{0}, \nonumber \\
	\hat{Z}(\theta) \ket{1} &\to e^{-i\frac{\theta}{2}}\ket{1}.
\end{align}

Now we employ the encoding for logical basis states,
\begin{align}
\ket{0}_L &\equiv \ket{0}_1\otimes\ket{1}_2,\nonumber \\
\ket{1}_L &\equiv \ket{1}_1\otimes\ket{0}_2,
\end{align}
Note that both logical basis states are invariant under a common $Z$-rotation (the other two physical basis states, $\ket{0}_1\otimes\ket{0}_2$ and $\ket{1}_1\otimes\ket{1}_2$ do not observe this property),
\begin{align}
	\hat{Z}_1(\theta)\hat{Z}_2(\theta)\ket{0}_L &= \ket{0}_L,\nonumber\\
	\hat{Z}_1(\theta)\hat{Z}_2(\theta)\ket{1}_L &= \ket{1}_L.
\end{align}
Thus, when acting on an arbitrary linear combination of these basis states (i.e a logical qubit),
\begin{align}
\ket\psi_L = \alpha \ket{0}_L + \beta\ket{1}_L,
\end{align}
via linearity the logical qubit must also be invariant under the collective error,
\begin{align}
	\hat{Z}_1(\theta)\hat{Z}_2(\theta)\ket\psi_L = \ket\psi_L.
\end{align}
This type of DFS encoding therefore protects a logical qubit against arbitrary, unknown but correlated $Z$-rotations.

The same principle can be logically extended to many other correlated error models. For example, operating in a rotated basis (under a Hadamard transform), one could similarly protect against correlated $X$-rotations using the encoding,
\begin{align}
\ket{0}_L &\equiv \ket{-}_1\otimes\ket{+}_2,\nonumber \\
\ket{1}_L &\equiv \ket{+}_1\otimes\ket{-}_2,
\end{align}
where \mbox{$\ket\pm=\frac{1}{\sqrt{2}}(\ket{0}\pm\ket{1})$}.

This idea, although very simple, is very powerful, since it is a completely passive form of error correction, requiring no syndrome measurements, feedforward or correction operations. It also arises quite naturally in some systems where undesired external fields act roughly uniformly across the physical qubits within a system.

@@

Let us illustrate how this can be achieved using a simple example, whereby a single-qubit system couples with an environment which introduces a slow, unknown phase evolution. Discretising time, we can write this phase evolution as a Pauli $Z$-rotation on the Bloch sphere\index{Bloch sphere},
\begin{align}
	\hat{Z}(\theta) = e^{i\frac{\theta}{2}\hat{Z}} = \left(\begin{matrix}
  e^{i\frac{\theta}{2}} & 0 \\
  0 & e^{-i\frac{\theta}{2}}
\end{matrix}\right),
\end{align}
for some unknown, but small $\theta$. Next we observe that, quite obviously, a $Z$-rotation of angle $\theta$ can be trivially undone by applying another $Z$-rotation of angle $-\theta$, since,
\begin{align}
\hat{Z}(-\theta)\hat{Z}(\theta) = \hat{\mathbb{I}}.	
\end{align}

@@

Now if we proceed for two time-steps, once bit-flipped and another not, we have,
\begin{align}
\underbrace{\hat{Z}(\theta)}_{\mathrm{forward}}\cdot\underbrace{[\hat{X}\hat{Z}(\theta)\hat{X}]}_{\mathrm{reverse}} = \hat{\mathbb{I}},
\end{align}
and the unknown phase-rotation has been eliminated.

The coupling to the external environment and the qubit state needn't be constant over time, but could be time-varying. In this instance, it is essential that the decoupling control sequence of bit-flip operations be much faster than the rate at which both the phase-rotations and computational operation are changing, such that consecutive unknown phase-rotations are approximately equal and thereby almost perfectly cancel. This rapid control sequence is sometimes referred to as `bang-bang' control\index{Bang-bang control}, since we are repeatedly implementing bit-flips at a fast rate.

@@

Dynamical decoupling extends to all manner of error models, beyond the simple $Z$-error model presented above. They are therefore a very powerful tool in error correction. However, unfortunately they are only naturally suited to continuous evolutions governed by Hamiltonians of the form shown in Eq.~(\ref{eq:dyn_dec_ham}), not to the more common discretised models such as the circuit and cluster state models.
